pdp: An R Package for Constructing Partial Dependence Plots

### 2017-01-02 manuscript/greenwell.tex
add ranger support

Currently supported models are described in Table~\ref{tab:models}. In these cases, \code{partial} should be able to automatically determine an appropriate value for \code{type} (\code{"regression"} or \code{"classification"}). In other situations, the user may need to specify this argument. This allows \code{partial} to be flexible enough to handle many of the model types not listed in Table~\ref{tab:models}; for example, neural networks from the \CRANpkg{nnet} package \citep{venables-modern-2002} and projection pursuit regression \citep{friedman-ppr-1981} using the \code{ppr} function in the \pkg{stats} package.

If \code{train = NULL}, \code{partial} tries to extract the original training data from the fitted model object. For objects that typically store a copy of the training data (e.g., objects of class \code{"BinaryTree"}, \code{"RandomForest"}, and \code{"train"}), this is straightforward. Otherwise, \code{partial} will attempt to extract the call stored in \code{object} (if available) and use that to evaluate the training data in the same environment from which \code{partial} was called. This can cause problems when, for example, the training data have been changed after fitting the model, but before calling \code{partial}. Hence, it is good practice to always supply the training data via the \code{train} argument in the call to \code{partial}\footnote{For brevity, we ignore this option in most of the examples in this paper.}.

\begin{table}[!htbp]
  \begin{tabular}{p{4cm}ll}
    \toprule
      Type of model & R package & Object class \\
      \midrule
      Decision tree             & \CRANpkg{C50} \citep{C50-pkg} & \code{"C5.0"} \\
                                & \pkg{party}    & \code{"BinaryTree"} \\
                                & \pkg{partykit} & \code{"constparty"}/\code{"party"} \\
                                & \CRANpkg{rpart} \citep{rpart-pkg} & \code{"rpart"} \\
      Bagged decision trees     & \CRANpkg{adabag} \citep{adabag-pkg} & \code{"bagging"} \\
                                & \CRANpkg{ipred} \citep{ipred-pkg} & \code{"classbagg"}, \code{"regbagg"} \\
      Boosted decision trees    & \CRANpkg{adabag} \citep{adabag-pkg} & \code{"boosting"} \\
                                & \pkg{gbm}      & \code{"gbm"} \\
                                & \CRANpkg{xgboost} & \code{"xgb.Booster"} \\
      Cubist                    & \CRANpkg{Cubist} \citep{Cubist-pkg} & \code{"cubist"} \\
      Random forest & \pkg{randomForest} & \code{"randomForest"} \\
      Conditional random forest & \pkg{party}    & \code{"RandomForest"} \\
                                & \pkg{partykit} & \code{"cforest"} \\
      Linear model              & \pkg{stats}    & \code{"lm"} \\
      Generalized linear model  & \pkg{stats}    & \code{"glm"}, \code{"lm"} \\
      Nonlinear least squares   & \pkg{stats}    & \code{"nls"} \\
      Multivariate adaptive regression splines (MARS) & \CRANpkg{earth} \citep{earth-pkg} & \code{"earth"} \\
      Support vector machine    & \CRANpkg{e1071} \citep{e1071-pkg} & \code{"svm"} \\
                                & \CRANpkg{kernlab} \citep{kernlab-pkg} & \code{"ksvm"} \\
      \bottomrule
  \end{tabular}
  \caption{Models specifically supported by the \pkg{pdp} package. \strong{Note:} for some of these cases, the user may still need to supply additional arguments in the call to \code{partial}.}
  \label{tab:models}
\end{table}

When additional arguments are necessary, the user will be prompted with an informative error message. For instance, the user may occasionally see the following:
\begin{example}
Error: The training data could not be extracted from object. Please supply
the raw training data using the `train` argument in the call to `partial`.
\end{example}
This will occur, for example, when using partial with the \code{"xgb.Booster"} class (see Section~\ref{sec:xgboost}). In fact, using \code{partial} with any object that does not store the original training data will result in the above error message.

>>>>

Currently supported models are described in Table~\ref{tab:models}. In these cases, \code{partial} should be able to automatically determine an appropriate value for the argument \code{type} (i.e., \code{"regression"} or \code{"classification"}). In other situations, the user may need to specify this argument. This allows \code{partial} to be flexible enough to handle many of the model types not listed in Table~\ref{tab:models}; for example, neural networks from the \CRANpkg{nnet} package \citep{venables-modern-2002} and projection pursuit regression \citep{friedman-ppr-1981} using the \code{ppr} function in the \pkg{stats} package.

\begin{table}[!htbp]
  \begin{tabular}{p{4cm}ll}
    \toprule
      Type of model & R package & Object class \\
      \midrule
      Decision tree             & \CRANpkg{C50} \citep{C50-pkg} & \code{"C5.0"} \\
                                & \pkg{party}    & \code{"BinaryTree"} \\
                                & \pkg{partykit} & \code{"constparty"}/\code{"party"} \\
                                & \CRANpkg{rpart} \citep{rpart-pkg} & \code{"rpart"} \\
      Bagged decision trees     & \CRANpkg{adabag} \citep{adabag-pkg} & \code{"bagging"} \\
                                & \CRANpkg{ipred} \citep{ipred-pkg} & \code{"classbagg"}, \code{"regbagg"} \\
      Boosted decision trees    & \CRANpkg{adabag} \citep{adabag-pkg} & \code{"boosting"} \\
                                & \pkg{gbm}      & \code{"gbm"} \\
                                & \CRANpkg{xgboost} & \code{"xgb.Booster"} \\
      Cubist                    & \CRANpkg{Cubist} \citep{Cubist-pkg} & \code{"cubist"} \\
      Random forest             & \pkg{randomForest} & \code{"randomForest"} \\
                                & \pkg{party}        & \code{"RandomForest"} \\
                                & \pkg{partykit} & \code{"cforest"} \\
                                & \CRANpkg{ranger} \citep{ranger-pkg} & \code{"ranger"} \\

      Linear model              & \pkg{stats}    & \code{"lm"} \\
      Generalized linear model  & \pkg{stats}    & \code{"glm"}, \code{"lm"} \\
      Nonlinear least squares   & \pkg{stats}    & \code{"nls"} \\
      Multivariate adaptive regression splines (MARS) & \CRANpkg{earth} \citep{earth-pkg} & \code{"earth"} \\
      Support vector machine    & \CRANpkg{e1071} \citep{e1071-pkg} & \code{"svm"} \\
                                & \CRANpkg{kernlab} \citep{kernlab-pkg} & \code{"ksvm"} \\
      \bottomrule
  \end{tabular}
  \caption{Models specifically supported by the \pkg{pdp} package. \strong{Note:} for some of these cases, the user may still need to supply additional arguments in the call to \code{partial}.}
  \label{tab:models}
\end{table}

Another important argument is \code{train}. If \code{train = NULL} (the default), \code{partial} tries to extract the original training data from the fitted model object. For objects that typically store a copy of the training data (e.g., objects of class \code{"BinaryTree"}, \code{"RandomForest"}, and \code{"train"}), this is straightforward. Otherwise, \code{partial} will attempt to extract the call stored in \code{object} (if available) and use that to evaluate the training data in the same environment from which \code{partial} was called. This can cause problems when, for example, the training data have been changed after fitting the model, but before calling \code{partial}. Hence, it is good practice to always supply the training data via the \code{train} argument in the call to \code{partial}\footnote{For brevity, we ignore this option in most of the examples in this paper.}. If \code{train = NULL} and the training data can not be extracted from the fitted model, the user will be prompted with an informative error message (this will occur, for example, when using \code{partial} with \code{"ksvm"} and \code{"xgb.Booster"} objects):
\begin{example}
Error: The training data could not be extracted from object. Please supply
the raw training data using the `train` argument in the call to `partial`.
\end{example}
